---
title: "Data Science in a Box Lab 4"
author: "Alex Lindquist, Nate Smullen, Wynter Malone, Yadhira Garcia"
date: "10/17/2022"
output: html_notebook
---

##Load Packages

```{r}
library(tidyverse)
library(dsbox)
```

##Alaska
```{r}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```
There are 3 Denny's in Alaska.

```{r}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
There are 2 LaQuintas in Alaska.

```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```
There are 6 observations in this joined dataframe, as there are 3 Denny's and 2
LaQuintas, so there are 6 total combinations of the two. The Denny's are 
represented with .x variable names whereas the LaQuintas are represented with 
the .y variable names.

```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

```{r}
dn_lq_ak <- dn_lq_ak %>% mutate(distance = haversine(longitude.x, latitude.x, 
                                                     longitude.y, latitude.y))
```

```{r}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
summary(dn_lq_ak_mindist)
dn_lq_ak_mindist %>% ggplot(aes(x=closest)) + geom_histogram()
```
There are only 3 data points, so it is difficult to describe the distribution.
All of the Denny's have a LaQuinta within 6km.

##North Carolina 
```{r}
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)
```
There are 28 Denny's in North Carolina.

```{r}
lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)
```
There are 12 LaQuintas in North Carolina.

```{r}
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc
```
There are 336 observations in this joined dataframe, as there are 28 Denny's and 
12 LaQuintas. The Denny's are represented with .x variable names whereas the 
LaQuintas are represented with the .y variable names.


```{r}
dn_lq_nc <- dn_lq_nc %>% mutate(distance = haversine(longitude.x, latitude.x, 
                                                     longitude.y, latitude.y))
```

```{r}
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
summary(dn_lq_nc_mindist)
dn_lq_nc_mindist %>% ggplot(aes(x=closest)) + geom_histogram()
```
The distribution for the closest LaQuinta to each Denny's is skewed right.

##Texas
```{r}
dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)
```
There are 200 Denny's in Texas.

```{r}
lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(lq_tx)
```
There are 237 LaQuintas in Texas.

```{r}
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx
```
There are 47,400 observations in this joined dataframe, as there are 200 Denny's
and 237 LaQuintas. The Denny's are represented with .x variable names whereas
the LaQuintas are represented with the .y variable names.


```{r}
dn_lq_tx <- dn_lq_tx %>% mutate(distance = haversine(longitude.x, latitude.x, 
                                                     longitude.y, latitude.y))
```

```{r}
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
summary(dn_lq_tx_mindist)
dn_lq_tx_mindist %>% ggplot(aes(x=closest)) + geom_histogram()
```
The distribution for the closest LaQuinta to each Denny's is skewed right, with
a majority of the Denny's having a Laquinta within 10 kilometers.
